---
title: "Use Filter To Identify Identical Elements of two Columns"
author: "Peter von Rohr"
date: "2021-02-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer
Tests and experiments on finding identical elements using filters


# Data
The data are used are given by certain pedigrees. The test pedigree is read from the following input file.

```{r}
vec_ped_path <- list.files(system.file('extdata', package = 'qprppedigree'), full.names = TRUE)
(s_ped_path <- vec_ped_path[4])
```

The pedigree is read using the following command

```{r}
tbl_p <- qprppedigree::read_prp_pedigree(ps_pedig_path = s_ped_path)
```

Records with identical IDs of animals and parents are to be filtered out. Start with the comparison of animal and mother

```{r}
require(dplyr)
sym_id <- sym("#IDTier")
sym_dam <- sym("IDMutter")
tbl_p %>% filter(!!sym_id == !!sym_dam)
```

The same can be done with the animal and the sire.

```{r}
tbl_p %>% filter(`#IDTier` == IDVater)
```

