---
title: "Programming with dplyr"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Disclaimer


# Background
When using tidy evaluations and other features of the [tidyverse](https://www.tidyverse.org/) in functions, special measures must be applied.


# Examples
From https://shipt.tech/https-shipt-tech-advanced-programming-and-non-standard-evaluation-with-dplyr-e043f89deb3d the following examples are tried. 

The idea is to use the following statement in a function. 

```{r}
library(dplyr)

mtcars %>%
  filter(cyl != 4) %>%
  group_by(gear) %>%
  summarize(avg_mpg = mean(mpg))

```

The first idea could look as follows

```{r}
# This won't work
test_function <- function(df, group_col){
  df %>%
    group_by(group_col) %>%
    summarize(n = n())
}

test_function(
  df        = mtcars, 
  group_col = gear
) 

```

But this does not work. 

The following solutions has to be used instead

```{r}
# We have a field name stored as text
my_field <- "cyl"

# Change our field name to a symbol object
my_symbol <- sym(my_field)

# Use it in a group_by call
mtcars %>%
  group_by(!!my_symbol) %>%
  summarize(mean_mpg = mean(mpg))
```

This can be used in a function as

```{r}
test_function_group_sym <- function(df, ps_group_field){
  # convert group field from string to symbol
  group_symbol <- sym(ps_group_field)
  # make the grouped-summarise
  df %>% 
    group_by((!!group_symbol)) %>%
    summarize(mean_mpg = mean(mpg))
}
test_function_group_sym(df = mtcars, ps_group_field = 'cyl')
```

The variable in the summarize-function can also be passed as in

```{r}
test_function_group_var_sym <- function(df, ps_group_field, ps_var_field, ps_var_name){
  # convert group field from string to symbol
  group_symbol <- sym(ps_group_field)
  # variable field
  var_symbol <- sym(ps_var_field)
  # summarise
  df %>%
    group_by((!!group_symbol)) %>%
    summarize(!!ps_var_name := mean((!!var_symbol)))
}
```
This can be tested with 'mpg'

```{r}
test_function_group_var_sym(df = mtcars, ps_group_field = 'cyl', ps_var_field = 'mpg', ps_var_name = 'mean_mpg')
```

The same tests with 'hp'

```{r}
test_function_group_var_sym(df = mtcars, ps_group_field = 'cyl', ps_var_field = 'hp', ps_var_name = 'mean_hp')
```


