---
title: "Use *-join in Functions"
output: html_notebook
---

# Disclaimer
This contains a how-to for using dplyr *-joins in functions.


# Other Examples
From https://community.rstudio.com/t/using-rlang-to-specify-joining-variables-in-dplyr-join-functions/48601

```{r}
set.seed(123)
library(dplyr, warn.conflicts = FALSE)

# Using a character vector as input
foo <- function(data, x) {
  inner_join(
    group_by_at(data, x) %>% summarise(wt = mean(wt)),
    group_by_at(data, x) %>% summarise(n = n()),
    by = x
  )
}

foo(mtcars, c("am", "cyl"))
```

Modify the above example

```{r}
library(dplyr)
(tbl_mean_wt <- mtcars %>% group_by_at(c("am", "cyl")) %>% summarise(wt = mean(wt)))
```

The second tbl

```{r}
(tbl_count_wt <- mtcars %>% group_by_at(c("am", "cyl")) %>% summarise(n = n()))
```

Joining the two tibbles

```{r}
inner_join(tbl_mean_wt, tbl_count_wt, by = c("am", "cyl"))
```

The same thing in a function

```{r}
vec_by_vars <- c("am", "cyl")
foo_join <- function(ptbl1, ptbl2, pvec_by){
  inner_join(ptbl1, ptbl2, by = pvec_by)
}
foo_join(ptbl1 = tbl_mean_wt, ptbl2 = tbl_count_wt, pvec_by = vec_by_vars)
```


# Data
The data are read from an existing pedigree, then a reproducible example is 
created and stored in a local file. Hence the following preparation steps 
should only be run once.

```{r eval=TRUE}
(vec_ped_path <- list.files(system.file('extdata', package = 'qprppedigree'), full.names = TRUE))
s_ped_path <- vec_ped_path[3]
tbl_test_ped <- qprppedigree::read_prp_pedigree(ps_pedig_path = s_ped_path)
dim(tbl_test_ped)
```

The first few lines of the pedigree read is given by

```{r}
head(tbl_test_ped)
```

From the complete pedigree, we select a number of records to produce the small example.

```{r}
library(dplyr)
tbl_ped_small_ex <- tbl_test_ped %>% filter(`#IDTier` == 997953372 |
                                            `#IDTier` == 997438624 |
                                            `#IDTier` == 997959197 |
                                            `#IDTier` == 997959200 |
                                            `#IDTier` == 997958313 |
                                            `#IDTier` == 997961951 |
                                            `#IDTier` == 997977775 |
                                            `#IDTier` == 998002894 |
                                            `#IDTier` == 998003890 |
                                            `#IDTier` == 998003889 |
                                            `#IDTier` == 997535439 |
                                            `#IDTier` == 997774809 |
                                            `#IDTier` == 997415332 |
                                            `#IDTier` == 997406925 |
                                            `#IDTier` == 1000810031 |
                                            `#IDTier` == 1000050624 |
                                            `#IDTier` == 999870937 |
                                            `#IDTier` == 1000161006)
```


The first few lines of the small pedigree consists of

```{r}
head(tbl_ped_small_ex)
```

The last few records are given by

```{r}
tail(tbl_ped_small_ex)
```

The small pedigree is written to a file to be able to read from that at a later time.

```{r}
s_small_ped_path <- 'small_pedigree.csv'
readr::write_delim(tbl_ped_small_ex, path = s_small_ped_path, delim = '|')
```


# Tests with Function
The tests starts with a fresh environment and then reads the data from the file

```{r}
rm(list = ls())
# path to test pedigree
s_small_ped_path <- 'small_pedigree.csv'
tbl_small_test_ped <- qprppedigree::read_prp_pedigree(ps_pedig_path = s_small_ped_path)
dim(tbl_small_test_ped)
```

The first few records of this pedigree corresponds to 

```{r}
head(tbl_small_test_ped)
```


## Birthdate Check
Based on this testpedigree, we want to check whether the birthdates between parents and offspring are consistent. Ultimately, this should be done in a function. We first start outside of a function.

The first step is to join the birthdates of the parents to the information about the animals.

```{r}
library(dplyr)
tbl_animal_bd <- tbl_small_test_ped %>% select(`#IDTier`, IDVater, Birthdate)
head(tbl_animal_bd)
```

The birthdates of the sires are joined from the original tibble

```{r}
tbl_animal_bd %>% 
  inner_join(tbl_small_test_ped, by = c('IDVater' = '#IDTier')) %>% 
  select(`#IDTier`, IDVater, Birthdate.x, Birthdate.y)
```

From the above tibble, we want to filter all records where the birthdate ('Birthdate.x') of the animal is before the birthdate ('Birthdate.y') of the sire. This is done using an additional filter-statement.

```{r}
tbl_animal_bd %>% 
  inner_join(tbl_small_test_ped, by = c('IDVater' = '#IDTier'), suffix = c(".Tier", ".Vater")) %>% 
  select(`#IDTier`, IDVater, Birthdate.Tier, Birthdate.Vater) %>%
  filter(Birthdate.Tier < Birthdate.Vater)
```

An alternative statement of the 'by'-argument

```{r}
pvec_by  <- c('IDVater' = '#IDTier')
tbl_animal_bd %>% 
  inner_join(tbl_small_test_ped, by = pvec_by, suffix = c(".Tier", ".Vater")) %>% 
  select(`#IDTier`, IDVater, Birthdate.Tier, Birthdate.Vater) %>%
  filter(Birthdate.Tier < Birthdate.Vater)

```

In order to be able to run the above statements in a function, we have to replace all column names in variables.

```{r}
ps_col_animal_id <- '#IDTier'
ps_col_sire_id <- 'IDVater'
ps_col_animal_bd <- 'Birthdate.Tier'
ps_col_parent_bd <- 'Birthdate.Vater'
# vectors 
pvec_by  <- c('IDVater' = '#IDTier')
pvec_suffix <- c(".Tier", ".Vater")

# using symbols
tbl_animal_bd %>% 
  inner_join(tbl_small_test_ped, by = pvec_by, suffix = pvec_suffix) %>% 
  select(!!sym(ps_col_animal_id), !!sym(ps_col_sire_id), !!sym(ps_col_animal_bd), !!sym(ps_col_parent_bd)) %>%
  filter(!!sym(ps_col_animal_bd) < !!sym(ps_col_parent_bd))

```

Convert the above statements into a function leads to

```{r}
check_parent_animal_bd <- function(ptbl_ped,
                                   ps_animal_id = '#IDTier',
                                   ps_parent_id,
                                   ps_col_bd    = 'Birthdate',
                                   pvec_by,
                                   pvec_suffix){
  # symbols
  sym_animal_id <- sym(ps_animal_id)
  sym_parent_id <- sym(ps_parent_id)
  sym_col_bd <- sym(ps_col_bd)
  vec_col_bd <- sapply(pvec_suffix, function(x) paste(ps_col_bd, x, sep = ''), USE.NAMES = FALSE)
  sym_animal_bd <- sym(vec_col_bd[1])
  sym_parent_bd <- sym(vec_col_bd[2])
  # create the tibble with animal and parent bd
  tbl_animal_bd <- dplyr::select(ptbl_ped, !!sym_animal_id, !!sym_parent_id, !!sym_col_bd)
  # join parent bd and run check with filter
  return(tbl_animal_bd %>%
           inner_join(ptbl_ped, by = pvec_by, suffix = pvec_suffix) %>% 
           select(!!sym_animal_id, !!sym_parent_id, !!sym_animal_bd, !!sym_parent_bd) %>%
           filter(!!sym_animal_bd < !!sym_parent_bd)) 
}
```

The function is tested with

```{r}
check_parent_animal_bd(ptbl_ped     = tbl_small_test_ped,
                       ps_parent_id = 'IDVater',
                       pvec_by      = c('IDVater' = '#IDTier'),
                       pvec_suffix  = c(".Tier", ".Vater"))
```

The same for the mothers

```{r}
check_parent_animal_bd(ptbl_ped     = tbl_small_test_ped,
                       ps_parent_id = 'IDMutter',
                       pvec_by      = c('IDMutter' = '#IDTier'),
                       pvec_suffix  = c(".Tier", ".Mutter"))
```

